---
title: "Simulation and Data Distributions"
output: html_notebook
---

## Simulation
It is the process of generation of data for a given purpose based on a choosen mathematical model. Simulation may be used produce datasets that are similar to data in another sample space or to reproduce datasets from random number generators.

## Random Number Generator
One of the most difficult tasks in computer science is to generate a random number that is completely unique. To avoid such tradeoffs, statiscal libraries as well as regular programming languages use pseudo-random number generators. Such pseudo-random numbers work on the concept of `seed`.

A `seed` is a set value that is used to reproduce datasets when using pseudo random number generators.
```{r}
SEED = 10
set.seed(SEED)
```

```{r}
set.seed(SEED)
x <- runif(10, 1, 100)
print('10 random numbers between 1-100')
x
```

All r\*, p\*, functions stated beneath behave in accordance with the seed that is set.

## Histogram and Density
Histogram is a simple way of representing the relationship between number of samples and the actual value for any given sample space. `hist()` function is used in R to generate histograms. Illustrations have been provided beneath.

## Data Distribution
There are varied linear models that can be used to simulate data over specific distributions. Any distribution may be referred as a function that describes the relationship between observations in a sample space.

`y = b0 + b1 * x + E`

Distributions are often described in terms of their density or density functions, which are usually represented using known mathematical models

## Types of Probability Distribution
There are various probability distributions avaialable for analysis in R including:
- Normal Distribution (`norm`)
- Uniform Distribution (`unif`)
- Poisson Distribution (`pois`)
- Binomial Distribution (`binom`)

```{r}
set.seed(SEED)
n <- 10
rnorm(n, mean = 0, sd = 1)

set.seed(SEED)
n <- 10
runif(n, min = 0, max = 1)

set.seed(SEED)
lambda <- 5
rpois(n, lambda)

set.seed(SEED)
size <- 3
prob <- 0.5
rbinom(n, size, prob)
```

## Random Sampling
Random sampling is used to select a certain number of samples on a random basis from a given sample space. It may be used in sample spaces that are not numeric.

```{r}
x <- 1:10

print('Selecting 5 unique samples from 1:10')
sample(x, 5)

print('Selecting 10 samples from 1:10 with replacement')
sample(x, 20, replace = T)
```

```{r}
print('10 random letters')
sample(letters, 10)

print('selecting 20 vowels')
sample(
  c("a", "e", "i", "o", "u"), 20,
  replace = T
)
```


## Exercise
Examine the orchis2 data object. Here are column data frame with a response variable `flower` and predictor variable `site`. Produce a histogram for the `site: 'sprayed'`. Now overlay a density plot over it.

```{r}
orchis2 <- read.csv('orchis2.csv')
orchis2
```

Histogram of frequency distribution for `site: sprayed`
```{r}
sprayed <- subset(orchis2, orchis2$site == 'sprayed')
x <- sprayed$flower
hist(x)
```

Histogram of density distribution for `site: sprayed`
```{r}
hist(x, freq = F)
```

Histogram of density distribution for `site: sprayed` with overlay density plot.
```{r}
hist(x, freq = F)
dens_x <- density(x)
lines(dens_x, lty = 'dashed')
dens_x_rect <- density(x, k = 'rectangular')
lines(dens_x_rect, lty = 'dotted')
```

## Other Datasets
Other datasets available for public use that comes bundled with R using the `datasets` library can be found from the link beneath.

https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/00Index.html

An example of using the `mtcars` dataset.

```{r}
library(datasets)
head(mtcars)
summary(mtcars)
```
