---
title: 'Experiment 8: Linear Regression and Hypothesis Testing'
output:
  word_document: default
  html_notebook: default
---

## Linear Regression

Linear regression is a model that is used in supervised learning to predict responses. The model is typically trained by use of a dataset that has features vector (X) and responses vector (y) present.

Any linear regression after training may be represented using the form:
`y = b0 + b1 * X`
where b0, b1 are weights/parameters of the model obtained after learning from train data.

Linear regression model can be useful for predicting in case of datasets that possess linear relationship amongst variables. The degree of correlation can be used as a measure of linear relationships.

```{r}
library(datasets)
data("faithful")
str(faithful)
summary(faithful)
```

```{r}
head(faithful)
```

Plot between eruptions and waiting in faitful dataset.
```{r}
plot(eruptions ~ waiting, data = faithful) # same as plot(faithful$waiting, faithful$eruptions)
```

```{r}
cor(faithful$eruptions, faithful$waiting)
```
Depicts a strong linear relationship.

```{r}
lr_model <- lm(eruptions ~ waiting, data = faithful)
plot(eruptions ~ waiting, data = faithful)
abline(lr_model, col = 'red', lty = 'dashed')
```

The weights (b0, b1) determined after training phase are as follows:
```{r}
lr_model$coefficients
```

Predicting using linear regression learned model:

```{r}
y_pred_1 <- sum(lr_model$coefficients * c(1, 65))
y_pred_1 # predicted y for x = 65
```


## Hypothesis Testing

Hypothesis testing is used to test the hypothesis that we are trying to form on the basis of a given model.

```{r}
library(datasets)
data("mtcars")
```

Q1. Find out the `cyl` and `am` variable in `mtcars` dataset have a strong correlation (use chi square test) null hypothesis: they are correlated or not, find out.

```{r}
cars_cyl <- mtcars$cyl
cars_am <- mtcars$am
table(mtcars$cyl, mtcars$am)
#chisq.test(cars_cyl, cars_am)
```

Q2. Find out the relationship between `mpg` and `cyl`.

```{r}
cars_mpg <- mtcars$mpg
cars_cyl <- mtcars$cyl
cor(cars_mpg, cars_cyl)
```

Q3. Use T test to find the influence `mpg` and `am`.

```{r}
batch_at <- subset(mtcars, am == 0)$mpg
batch_at
batch_mt <- subset(mtcars, am == 1)$mpg
batch_mt

t.test(batch_at, batch_mt)
```

Use chisql.test(), t.test() functions.