---
title: 'Experiment 1: Basic Operations'
output:
  word_document: default
  html_notebook: default
---

## Variables

Variables are used to hold data values of any type.

### Creating Variables

```{r}
x <- 131198
y <- "Hello World"
print(x)
print(y)
```

### Modifying Values of Variables

```{r}
print(x) # initial value
x <- 2 * x # value of x will be doubled
print(x) # final value
```

### Removing Existing Variables

```{r}
rm(x)
rm(y)
print('x, y variables were removed')
```

## Environment

An environment can be used to store multiple variables that are to be used for a common purpose. It is like a wrapper for holding multiple values in named fashion.

### Creating an Environment

```{r}
anEnv <- new.env()
```

### Assign Variables to the Environment

```{r}
# assign a value to a variable
# inside created environment
assign('name', "Contoso", anEnv)
# another one
assign('age', 23, anEnv)
```

### Get the Value of the Variables

```{r}
# get the variable values
get('name', anEnv)
```

```{r}
get('age', anEnv)
```

### Display all Variables in Environment

```{r}
ls(anEnv)
```

### Remove Variables from Environment

```{r}
rm('name', envir = anEnv)
print("removed variable 'name' from anEnv")
```

```{r}
ls(anEnv) # only age will exist
```

### Remove the Environment

```{r}
rm(anEnv)
print('removed environment anEnv')
```

## Working Directory

Working directory refers to the current directory relative to which all script executions are held. The canonical alias of the standard unix-like command `pwd` is `getwd()` in R.

### Get Current Directory
```{r}
getwd()
```

### Set the Current Directory

```{r}
# it may be set either relative to current directory 
setwd('../CSE2019')
afterFirstSet <- getwd()
# or relative to system root
setwd('/Users/Shared')
afterSecondSet <- getwd()
```

```{r}
print(afterFirstSet)
print(afterSecondSet)
```

## Few Mathematical Functions

### Logarithm Function

```{r}
log(16, base = 4) # using with custom base
log2(8) # using with 2 as base
log10(10000) # using with 10 as base
log(exp(10), base = exp(1)) # natural logarithm, base = e
```

### Exponential Function

```{r}
# used to compute euler's const
# and higher powers of exp fn
exp(1) # euler's const
exp(10) # e ^ 10
```

### Absolute Value

```{r}
# similar to modulus
# computes absolute value of a numeric
abs(-909)
```

### Factorial

```{r}
factorial(5) # 5 * 4 * 3 * 2 * 1
```

### Square Root

```{r}
sqrt(144) # will give 12 as 12 ^ 2 = 144
```

## Data Types

### Numeric

```{r}
numY <- 17.29
numY
```

### Integer

```{r}
intX <- 69L # L is used to denote it's integer value
intX
```

### Complex

```{r}
point <- 3 + 2i
print(point)
```

### Logical

```{r}
flag <- TRUE
print(flag)
anotherFlag <- FALSE
print(anotherFlag)
```

### Character

```{r}
name <- "Contoso Smith"
print(name)
```

### Date and Time

```{r}
aDate <- Sys.Date() # current date
aTime <- Sys.time() # current time
print(aDate)
print(aTime)
```

### Coercion - Type Casting

```{r}
intY <- as.integer(9.9)
print(intY)
```

```{r}
numZ <- as.numeric(9L)
print(numZ)
```

```{r}
numA <- as.numeric(flag)
print(numA) # logical bool to numeric
```

### Determining the Data Type

```{r}
class(flag)
class(intY)
class(aTime)
class(aDate)
class(numY)

class(intX)
class(numZ)
```

### Type Tests

```{r}
is.integer(9)
is.character(intX)
is.character(name)
is.complex(point)
```

### Non Existent Values

```{r}
NaN
NA
1 / 0
```

```{r}
is.finite(1 / 0)
is.infinite(Inf)

as.logical(name) # will produce NA, since x is character

# example of NaN
print(Inf - Inf)
```

## Characters

A string, as in a sequence of characters belong to the class `character` in R.

### Showing Part of a Character Variable

```{r}
aStr <- "Hello World"
substr(aStr, 7, 11) # start stop index specified (inclusive)
```

### Concatenate Two Character Variable

```{r}
x <- "Welcome"
y <- "Change"
paste(x, y) # by default space used as seperator
```

### Length of Character Variable

```{r}
nchar(x)
```

The experiment was also designed to perform a set of simple calculations using the R language so as to make us familiar with basic R syntax.

7 such sample calculations have been performed.

## Number of Hours in a Leap Year

```{r}
# hours in leap year
leapYearHours <- 24 * 366
print(leapYearHours)
```

## Convert Kilometer to meter

```{r}
# convert km to m
km <- readline('Input km: ')
km <- as.numeric(km)
print(paste('Output: ', km * 1000, 'm'))
```

## Number of Hours in a Non-Leap Year

```{r}
# hours in a year
yearHours <- 24 * 365
print(yearHours)
```

## Find Square Root

```{r}
inputNumber <- as.numeric(readline('Input number: '))
print(paste("Square Root: ", sqrt(inputNumber)))
```

## Find the Volume of a Rod

The volume of a rod is measured using the following formula:
`V = pi * (r ^ 2) * h`

```{r}
# volume of rod
radius = as.numeric(readline("Radius input: "))
height = as.numeric(readline("Height input: "))
volume = pi * radius * radius * height
print(paste("Volume is ", volume))
```

## Value of a Light Year

Light year is a unit of measurement of distance defined by the distance travelled by light in 1 year.

```{r}
# value of light year
secondsInAYear = yearHours * 3600
speedOfLight = 3e8
lightYear = speedOfLight * secondsInAYear
print(paste(as.character(lightYear),  "m"))
```

## Time since Unix Epoch

Unix Epoch is defined to be the date 1 January, 1970. Most programming platforms allow use of datetime data type by internally storing the number of milliseconds elapsed since Unix Epoch. This approach expresses time as long values.

```{r}
# approx time since Unix epoch
# assuming all years have 365 days

# this is solely an approximation
# due to ease of calculations
yearsSinceEpoch = 2019 - 1970
print(yearsSinceEpoch * secondsInAYear * 1000)
```

```{r}
# using Unix Epoch long ms
timeInSeconds = as.numeric(Sys.Date())
timeInMs = timeInSeconds * 1000
print(timeInMs)
```


