---
title: "Experiment 2: Simple Calculations"
output: html_notebook
---

This experiment was designed to perform a set of simple calculations using the R language so as to make us familiar with basic R syntax.

7 such sample calculations have been performed.

## Number of Hours in a Leap Year

```{r}
# hours in leap year
leapYearHours <- 24 * 366
print(leapYearHours)
```

## Convert Kilometer to meter

```{r}
# convert km to m
km <- readline('Input km: ')
km <- as.numeric(km)
print(paste('Output: ', km * 1000, 'm'))
```

## Number of Hours in a Non-Leap Year

```{r}
# hours in a year
yearHours <- 24 * 365
print(yearHours)
```

## Find Square Root

```{r}
inputNumber <- as.numeric(readline('Input number: '))
print(paste("Square Root: ", sqrt(inputNumber)))
```

## Find the Volume of a Rod

The volume of a rod is measured using the following formula:
`V = pi * (r ^ 2) * h`

```{r}
# volume of rod
radius = as.numeric(readline("Radius input: "))
height = as.numeric(readline("Height input: "))
volume = pi * radius * radius * height
print(paste("Volume is ", volume))
```

## Value of a Light Year

Light year is a unit of measurement of distance defined by the distance travelled by light in 1 year.

```{r}
# value of light year
secondsInAYear = yearHours * 3600
speedOfLight = 3e8
lightYear = speedOfLight * secondsInAYear
print(paste(as.character(lightYear),  "m"))
```

## Time since Unix Epoch

Unix Epoch is defined to be the date 1 January, 1970. Most programming platforms allow use of datetime data type by internally storing the number of milliseconds elapsed since Unix Epoch. This approach expresses time as long values.

```{r}
# approx time since Unix epoch
# assuming all years have 365 days

# this is solely an approximation
# due to ease of calculations
yearsSinceEpoch = 2019 - 1970
print(yearsSinceEpoch * secondsInAYear * 1000)
```

```{r}
# using Unix Epoch long ms
timeInSeconds = as.numeric(Sys.Date())
timeInMs = timeInSeconds * 1000
print(timeInMs)
```