---
title: 'Experiment 3: Data Frames and Factors'
output:
  word_document: default
  html_notebook: default
---

## Data Frames

### Creating a Data Frame

```{r}
n <- 1:5
s <- rep('Name', 5)
x <- c(rep(TRUE, 3), rep(FALSE, 2))

df <- data.frame(n, s, x)
df
```

```{r}
class(df)
class(df[1])
class(df[[1]])
is.vector(df[[1]])
class(df$n)
```

### Naming columns 

```{r}
names(df) = c("Number", "Name", "Student")
df
```

### Strings As Factors

```{r}
class(df$Name)
df2 <- data.frame(n, s, x, stringsAsFactors = F)
class(df2$s)
```

### Create a sample data frame

```{r}
serialNum <- 1:10
names <- paste("Name", as.character(serialNum))
salary <- 33000 * serialNum
joiningDates <- rep(Sys.Date(), length(serialNum))

employees <- data.frame(
  serialNum, names, salary, joiningDates,
  stringsAsFactors = F
)
print(employees)
```

## Factors

### Sample factors

```{r}
gender <- factor(c("Male", "Female", "Others"))
class(gender)
directions <- c("East", "West", "North", "South")
print(directions)
print(is.factor(directions))

gender[[1]]
levels(gender) <- c("male", "female", "others")
gender
```

```{r}
print(as.factor(directions))
```

### Create a Data Frame and Check for factors

```{r}
height <- seq(30, 50, by = 1.8)
weight <- seq(40, 60, by = 1.8)
halfLength <- length(height) / 2
gender <- c(rep('Male', halfLength), rep('Female', halfLength))

bmi <- data.frame(height, weight, gender)
bmi
```

```{r}
is.factor(bmi$gender)
```

## Simple exercise on factors and dataframes

Execute the following, observe and interpret the output:
a <- c(1, -5, 3.4, -2, 5, 3)
factor(a)
factor(a, 1:5)
factor(a, 3:5)
factor(a, -10:5)

```{r}
a <- c(1, -5, 3.4, -2, 5, 3)
factor(a)
factor(a, 1:5)
factor(a, 3:5)
factor(a, -10:5)
```

Create the following vectors of size 10
v1: vector of first names
v2: vector of last names
v3: vector of random nos from 1 to 50

```{r}
v1 <- c(
  "John", "Joe", "Jacob", "Steve",
  "Candis", "Jenny", "Angel", "Jill", "Mary"
  )
v2 <- c(
  "Smith", "Allen", "Manor", "Gosling",
  "Pegasus", "Trakersky", "Archer", "Jeffry", "Kim"
  )
length(v1) == length(v2)
```

```{r}
# 20 random numbers between 1 to 50
v3 <- sample(
    seq(0, 50, by = 0.1),
    length(v1)
  )
v3
```

Create a datafram from these vectors with appropriate headings.

```{r}
people <- data.frame(v3, v1, v2, stringsAsFactors = F)
names(people) <- c("UniqueSerial", "FirstName", "LastName")
people
```

Add a column ContactNumber to this data frame.

```{r}
people[[4]] <- c(
  1e10 + seq(1, length(v1))
)
names(people)[4] <- "ContactNumber"
people
```

```{r}
column3 <- people[[1]]
class(column3)
column3 <- as.factor(column3)
people$UniqueSerial <- column3
class(people$UniqueSerial)
```


Create another df with columns:
enrolment_number, semester and merge them with existing df

```{r}
enN <- 17002020 + seq(1, length(v1))
sem <- sample(1:8, length(v1), replace = T)
sem
enN
```

```{r}
people <- cbind(people, enN, sem)
people
```

```{r}
names(people)[5] <- "EnrolmentNumber"
names(people)[6] <- "Semester"
people
```
