---
title: "Datasets and Simple Visualisations"
output:
  html_notebook: default
---

## Datasets in R

R provides us with a few standard datasets using the default library called `datasets`. It can be used as follows:

```{r}
library(datasets)
```

A variety of publicly available standard datasets are available as a part of the UCI Machine Learning Repository which can found at https://archive.ics.uci.edu/ml/datasets.html

Similarly, the R datasets library (https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/00Index.html) has the `iris` dataset.

The data is loaded and a summary of the dataset is generated.

```{r}
# load iris dataset
data("iris")
# generate a summary for it
summary(iris)
```

A few rows from the beginning of the dataset are diplayed.

```{r}
head(iris)
```

The structure of the `iris` dataset is examined.
```{r}
str(iris)
```

## Simple Visualisations

`plot()` function is used to generate graphs.

```{r}
# all possible plots from the dataframes
# using each pair of columns
plot(iris)
```

A plot between `iris` flowers' sepal lengths and petal lengths.

```{r}
x <- iris$Sepal.Length
y <- iris$Petal.Length
plot(x, y)
```

### Histograms

A histogram representing different sepal lengths.

```{r}
hist(iris$Sepal.Length)
```

### Subset

`subset()` is used to generate a data frame that consists of only those rows which satisfy a given condition.

```{r}
# data frame of only 
# iris setosa
ss <- subset(iris, Species == 'setosa')
# display only few elements from
# the subset, since it has lot of rows
head(ss)
```

### Box Plots

`boxplot()` to generate box plots.

```{r}
boxplot(iris)
```

`par()` can be used to set many graphical parameters eg. margin. 

```{r}
par(
  mar = c(5, 5, 3, 3) # bottom, left, top, right margins (in order)
)
boxplot(iris)
```

## Read and Write Datasets

Reading and writing datasets using the `read.*()` and `write.*()` family of functions.

```{r}
write.csv(iris,file="iris.csv")
```

```{r}
df1 <- read.csv("iris.csv")
head(df1)
```